Assessing United States county-level exposure for research on tropical cyclones and human health

### 2019-11-28 writing/anderson_ehp_supp.tex
Update figure references
Work on paper, use 34 knot threshold
Continue work on text

\affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
  State University, Fort Collins, CO, 80523} 
\affil[2]{University of Colorado Denver School of Medicine, Aurora, CO, 80045} 
\affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
  Center, Huntsville, AL, 35805}
\affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
  University, Fort Collins, CO, 80523} 
\affil[5]{Department of Industrial and Operations Engineering, University of 
  Michigan, Ann Arbor, MI, 48109}
\affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
\affil[7]{Debian and R Projects; Department of Statistics, University of
  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
\affil[8]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
  Health, Baltimore, MD, 21205}
\affil[*]{Corresponding author: Brooke Anderson, brooke.anderson@colostate.edu}

@@

@@

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{averageexposureonlysupp}
\caption{Average number of storm exposures per decade in U.S. counties for each 
	exposure metric, limited analysis to years for which data on all five 
	exposures were available (1996\,--\,2011). The criteria behind each of 
	the five metrics is given in Table 1 of the main text. 	
	} 
\label{fig:averageexposuresupp}
\end{figure}

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{windexamples}
\caption{Comparison of county-level estimates of maximum sustained surface wind speeds
	for Hurricane Ike in 2008 (top) and Hurriance Sandy in 2012 (bottom). 
	Each map shows the estimated maximum sustained surface wind speed 
	classification ($<$34~\si{\knot}; 34\,--\,49.9~\si{\knot}; 
	50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}) for each study county. 
	The maps labelled ``Modeled'' (left)
	shows the classifications based on modeled sustained surface wind
	speeds, which were included in the open-source data as the main 
	wind metric and used in further analysis in this research. The maps
	labelled ``Extended Best Tracks'' (right) show classifications based on the 
	wind radii given in \ac{HURDAT2} (included as a secondary wind 
	metric in the open-source data). The red lines show the storms' 
	tracks.
	}
\label{fig:windexamples}
\end{figure}

>>>>

\affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
  State University, Fort Collins, CO, 80523} 
\affil[2]{University of Colorado Anschutz Medical Campus, School of Medicine, Aurora, CO, 80045} 
\affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
  Center, Huntsville, AL, 35805}
\affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
  University, Fort Collins, CO, 80523} 
\affil[5]{Department of Industrial and Operations Engineering, University of 
  Michigan, Ann Arbor, MI, 48109}
\affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
\affil[7]{Department of Statistics, University of
  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
\affil[8]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
  Health, Baltimore, MD, 21205}
\affil[*]{Corresponding author: Brooke Anderson, brooke.anderson@colostate.edu}

@@

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{windexamples}
\caption{Comparison of county-level estimates of peak sustained surface wind 
	for Hurricane Ike in 2008 (top) and Hurriance Sandy in 2012 (bottom). 
	Each map shows the estimated peak sustained surface wind 
	classification ($<$34~kt; 34\,--\,49.9~kt; 
	50\,--\,63.9~kt; $\ge$64~kt) for each study county. 
	The maps labelled ``Modeled'' (left)
	shows the classifications based on modeled peak sustained surface wind,
	which were included in the open-source data as the main 
	wind metric and used in further analysis in this research. The maps
	labelled ``Extended Best Tracks'' (right) show classifications based on the 
	wind radii given in \ac{HURDAT2} (included as a secondary wind 
	metric in the open-source data). The red lines show the storms' 
	tracks.
	}
\label{fig:windexamples}
\end{figure}

\begin{figure}[tbhp!]
\centering
\includegraphics[width=7cm]{averageexposureonlysupp}
\caption{Average number of storm exposures per decade in U.S. counties for each 
	single-hazard exposure metric, limited analysis to years for which data on all five 
	exposures were available (1996\,--\,2011). The criteria behind each of 
	the five metrics is given in Table 1 of the main text. 	
	} 
\label{fig:averageexposuresupp}
\end{figure}

@@

### 2019-11-28 writing/ehp_abstract.tex
Continue work on text

\noindent \textbf{Background:} 
        Tropical cyclone epidemiology can be 
	advanced through county-level exposure assessment methods that are comprehensive and 
	consistent across space and time, allowing multi-year, multi-storm studies. 
	Further, an understanding of 
	patterns in and between hazard-specific exposure metrics can help in 
	designing tropical cyclone epidemiological studies.\\ 
\textbf{Objectives:} (1)~Provide an open-source dataset for tropical
	cyclone exposure assessment for epidemiological research and
	(2)~investigate patterns and agreement between county-level assessments of tropical cyclone
	exposure based on different storm hazards. \\ 
\textbf{Methods:} We created an open-source dataset with county-level data on
	exposure to four tropical cyclone hazards: maximum sustained winds,
	rainfall, flooding, and tornadoes. The data cover all eastern \ac{US} counties 
	for all land-falling or near-land
	Atlantic basin storms for~1996\,--\,2011 for all
	metrics and up to~1988\,--\,2015 for specific metrics. 	
	We validated 
	measurements against other data sources and investigated patterns
	and agreement among binary exposure classifications based on these
	metrics, as well as between distance from the storm's track as a proxy for
	exposure.\\ 
\textbf{Results:} Measurements in the dataset created here typically
        agreed with data from other sources, and we present and discuss areas of
	disagreement and other caveats for the data.
        Long-term geographic patterns in exposure classification agree with previous results from
	atmospheric science. County-level exposure classification for a storm differs substantially 
	depending on the hazard metric used. \\ 
\textbf{Discussion:} Our results provide a multi-hazard dataset that can be
	leveraged for epidemiologic research on tropical cyclones, as well as 
	insights that can inform the design of tropical cyclone epidemiological 
	studies.
\end{abstract}

>>>>

\noindent \textbf{Background:} 
        Tropical cyclone epidemiology can be 
	advanced through county-level exposure assessment methods that are comprehensive and 
	consistent across space and time, facilitating multi-year, multi-storm studies. 
	Further, an understanding of 
	patterns in and between exposure metrics that are based on specific hazards of the storm can help in 
	designing tropical cyclone epidemiological studies.\\ 
\textbf{Objectives:} (1)~Provide an open-source dataset for tropical
	cyclone exposure assessment for epidemiological research and
	(2)~investigate patterns and agreement between county-level assessments of tropical cyclone
	exposure based on different storm hazards. \\ 
\textbf{Methods:} We created an open-source dataset with county-level data on
	exposure to four tropical cyclone hazards: peak sustained wind,
	rainfall, flooding, and tornadoes. The data cover all eastern \ac{US} counties 
	for all land-falling or near-land
	Atlantic basin storms, covering~1996\,--\,2011 for all
	metrics and up to~1988\,--\,2015 for specific metrics. 	
	We validated 
	measurements against other data sources and investigated patterns
	and agreement among binary exposure classifications based on these
	metrics, as well as compared them to use of distance from the storm's track as a proxy for
	exposure.\\ 
\textbf{Results:} We describe how we developed this dataset and explore how its
	measurements compare to other data that could be used to characterize
	tropical cyclone hazards for epidemiological research. Measurements in
	the dataset created here typically agreed with data from other sources,
	and we present and discuss areas of disagreement and other caveats for
	the data. Tropical cyclones typically brought different hazards to different
	counties, and so agreement was usually low between distance-based tropical
	cyclone exposure assessment and each hazard-specific metric, as well as 
	between pairs of hazard-specific metrics. \\
\textbf{Discussion:} Our results provide a multi-hazard dataset that can be
	leveraged for epidemiologic research on tropical cyclones, as well as 
	insights that can inform the design of tropical cyclone epidemiological 
	studies.
\end{abstract}

### 2019-11-28 writing/ehp_cover.tex
Rearrange content
Continue work on text

\author[1,*]{G. Brooke Anderson} 
\author[1,2]{Joshua Ferreri}
\author[3]{Mohammad Al-Hamdan} 
\author[3]{William Crosson} 
\author[4]{Andrea Schumacher} 
\author[5]{Seth Guikema} 
\author[6]{Steven Quiring} 
\author[7]{Dirk Eddelbuettel} 
\author[1]{Meilin Yan} 
\author[8]{Roger D. Peng}

\affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
  State University, Fort Collins, CO, 80523} 
\affil[2]{University of Colorado Denver School of Medicine, Aurora, CO, 80045} 
\affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
  Center, Huntsville, AL, 35805}
\affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
  University, Fort Collins, CO, 80523} 
\affil[5]{Department of Industrial and Operations Engineering, University of 
  Michigan, Ann Arbor, MI, 48109}
\affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
\affil[7]{Debian and R Projects; Department of Statistics, University of
  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
\affil[8]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
  Health, Baltimore, MD, 21205}
\affil[*]{Corresponding author: Brooke Anderson, brooke.anderson@colostate.edu,
  1681 Campus Delivery, Fort Collins, CO, 80523}

>>>>

\author[1,*]{G.~Brooke Anderson} 
\author[1,2]{Joshua Ferreri}
\author[3]{Mohammad Al-Hamdan} 
\author[3]{William Crosson} 
\author[4]{Andrea Schumacher} 
\author[5]{Seth Guikema} 
\author[6]{Steven Quiring} 
\author[7]{Dirk Eddelbuettel} 
\author[1]{Meilin Yan} 
\author[8]{Roger D.~Peng}

\affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
  State University, Fort Collins, CO, 80523} 
\affil[2]{University of Colorado Anschutz Medical Campus, School of Medicine, Aurora, CO, 80045} 
\affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
  Center, Huntsville, AL, 35805}
\affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
  University, Fort Collins, CO, 80523} 
\affil[5]{Department of Industrial and Operations Engineering, University of 
  Michigan, Ann Arbor, MI, 48109}
\affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
\affil[7]{Department of Statistics, University of
  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
\affil[8]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
  Health, Baltimore, MD, 21205}
\affil[*]{Corresponding author: G.~Brooke Anderson, brooke.anderson@colostate.edu,
  1681 Campus Delivery, Fort Collins, CO, 80523}

### 2019-11-28 writing/ehp_discussion.tex
Continue work on text

Epidemiological studies can help characterize which health risks are elevated
during disasters, to what degree, and to
whom~\parencite{ibrahim2005unfortunate, noji2005disasters}.  As a result, these
studies help improve disaster preparedness and
response~\parencite{noji2005disasters}.  However, tropical cyclones are
multi-hazard events, making it complicated to measure exposure and so to
conduct multi-community, multi-year studies leveraging large administrative
datasets.  Here, we provide open-source county-level data on multiple tropical
cyclone exposure metrics and explore limitations in that data.  
Further, we explore patterns in storm exposure classifications based on
different metrics, and we find that county-level tropical cyclone exposure
assessments vary substantially when using different metrics, based either on
storm hazards or on how close the storm's central track came to the community.
Our results can inform exposure assessment for future county-level studies of
the health risks and impacts associated with tropical cyclones exposure as
well as insights to inform epidemiological study design.

@@

We investigated how well the open-source data correspond with data from other
potential sources
(Figures~\ref{fig:raincomparison}\,--\,\ref{fig:floodcomparison} and~S2). While
generally in agreement, there are some caveats. The rainfall data are generally
well-correlated with ground-based observations, but may sometimes oversmooth
very high rainfall values compared to ground-based observations
(Figure~\ref{fig:raincomparison}). When rainfall data is used to create binary
exposure classifications, this disagreement is unlikely to influence results,
as both data sources agree in identifying these as storms with high rainfall,
but would be important to consider for cases that include rainfall as a
continuous measurement.

The wind data are based on modeled, rather than observed, values, and while the
modeled wind data generally agree well with post-analysis maximum wind radii
from \ac{HURDAT2}~\parencite{landsea2013} (Figure~\ref{fig:windcomparison}),
there were a few storms with some discrepancies. These storms---for example, Hurricane Sandy
in 2012 and Hurricane Ike in 2008---were
unusually large systems for which high winds persisted well inland from
landfall (Figures~\ref{fig:windcomparison}
and~S2).

For the flooding data, we found that flood event status determined based on the
NOAA Storm Events listings typically agreed with measurements from \ac{USGS}
streamgages, with a flood event more likely to be listed if a storm elevated
streamflow across the county (Figure~\ref{fig:floodcomparison}).  However,
there are differences between the two flooding datasets, and these highlight
both the difficulty of measuring flood exposure at the county level and
inherent challenges in using data from a storm event database for epidemiologic
exposure assessment.  For example, there was one storm in Fulton County, GA,
for which there was high streamflow but not an associated flood event listing
(Hurricane Ida, 2009).  This storm occurred in November 2009, following a month
with historic rainfall and flooding in
Georgia~\parencite{shepherd2011overview}.  In this case, the flooding
associated with Ida was incorporated into an ongoing flood event listing, with
a start date well before the five-day window we used to temporally match storm
event listings with tropical cyclone tracks.

@@

While these are important caveats for the data, we selected these data sources
as among the best currently available for measuring each of these hazards
consistently and comprehensively at a multi-county, multi-year scale. In
addition to providing tropical cyclone exposure metrics for individual hazards,
this dataset and its associated software allow users not only to access
measurements for single hazards, but also to create tropical cyclone exposure
profiles based on multiple hazards or to craft exposure indices that combine
hazard metrics~\parencite{chakraborty2005population, peduzzi2009assessing}.
This ability can be critical, as different hazards of tropical cyclones can
act synergistically in causing impacts~\parencite{smith2009}.

@@

We found that tropical cyclones tended to bring different hazards to different
counties, and so agreement was typically low between distance-based
tropical cyclone exposure assessment and each of the single hazard-specific
exposures, as well as between pairs of hazard-specific metrics
(Figures~\ref{fig:ivanexposure}\,--\,\ref{fig:jaccard} and S4\,--\,S5).  These
findings align with previous results from atmospheric science and related
fields on the characteristics of tropical cyclones. While tropical cyclone
rainfall and windspeed can be well-correlated when the tropical cyclone is over
water~\parencite{cerveny2000}, this relationship often weakens once the
hurricane has made landfall~\parencite{jiang2008}.  Fast-moving tropical
cyclones bring higher risks of dangerous winds inland~\parencite{kruk2010},
while slow-moving tropical cyclones are likely to bring more
rain~\parencite{rappaport2000} and cause more damage because of sustained
hazardous conditions~\parencite{rezapour2014}. Further, while the likelihood
and extent of flooding during a tropical cyclone is related to the tropical
cyclone's rainfall, it is also driven by factors like top soil saturation and
the structure of the water basin's drainage network~\parencite{chen2015,
rees2001}.

Therefore, the use of a distance-based metric to assess exposure to any of
these hazards, or the use of measurements from one hazard as a proxy for
exposure to any of the other hazards considered, would often introduce exposure
misclassification. This conclusion reinforces similar findings from a study of
Florida's 2004 storm season~\parencite{grabich2015measuring}.  For some
studies, such exposure misclassification might plausibly be differential.  For
example, tropical cyclone wind exposures tend to be concentrated in counties
near the coast, while tropical cyclone wind exposures sometimes extend well
inland.  If the etiologically-relevant exposure for a health outcome is extreme
rainfall but exposure is classified based on measurements of wind, the
probability of being misclassified as unexposed would be higher in inland
counties, while the probability of being misclassified as exposed would be
higher in coastal counties. If coastal counties differ from inland counties in
either the outcome of interest or in factors associated with risk of that
outcome, exposure misclassification would be
differential~\parencite{savitz2016interpreting}.  Such differential exposure
misclassification could bias estimates of tropical cyclone effects either
towards the null (estimating a lower or null association compared to the true
association that exists) or away from the null (estimating a larger association
than actually exists)~\parencite{savitz2016interpreting, armstrong1998effect}.

We did, however, find a small set of tropical cyclones for which for
which agreement was high across several single-hazard exposure assessments
(e.g., Floyd in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996; Ernesto
in~2006 (Figure~\ref{fig:jaccard})).  Hurricanes Floyd in~1999 and
Irene in~2011 both made their first \ac{US} landfall in North Carolina at minor
hurricane windspeeds (Category~2 and~1, respectively) and then skimmed the
eastern coast of the \ac{US} north through New England, bringing substantial
rainfall to much of the eastern coast from North Carolina north and causing
extensive inland flooding in North Carolina (Floyd) and New England
(Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.  Hurricanes Hannah
in~2008, Bertha in~1996, and Ernesto in~2006 also followed the eastern
coastline. For these storms, the tropical cyclones' persistent proximity
to water may have helped maintain wind speeds in similar patterns to rain and
distance-based exposure assessments, resulting in more similarities across
exposure assessments compared to the other tropical cyclones we considered.
For these storms, it may be possible to assess exposure to multiple hazards of
the storm using a single metric, in some cases even a proxy like the distance
between a county and the storm's track.  However, for these storms it may be
difficult to untangle the contribution of each hazard to the overall
effect of the storm, given that several hazards have similar geographical
patterns.

@@

The dataset presented here does have several limitations, in addition to the
caveats previously discussed. First, the dataset is not comprehensive of all
possible tropical cyclone hazards. For example, coastal counties can experience
dangerous storm surge, which is not specifically covered in this dataset
(although some resulting coastal flooding is captured). We are exploring ways
to include this in future versions of the dataset, although to date we have
focused on exposures that could affect any county, whether inland or coastal.
Second, this data is aggregated to the county level. Many health outcome
datasets are aggregated at this level, but some may be aggregated at a finer
spatial resolution (e.g., Census tract or ZIP code) or unaggregated (i.e.,
point locations for each outcome). We have published the wind model used to
create this dataset as its own open-source R package
\parencite{stormwindmodel}, and it can be used to model storm-associated winds
at a finer spatial resolution; however, measurements of other hazards cannot
similarly be re-scaled through tools we provide. Next, we provide these data
and associated software tools through R packages, and so some experience in the
R programming language is required to make full use of them. However, R is
currently a popular programming language for environmental epidemiology,
allowing the data to reach a large audience, and we are exploring options to
create a web application using the Shiny platform to allow broader web-based
access of the data \parencite{shiny2019}.  Finally, to assess patterns and
agreement for binary exposure classifications, we have chosen one set of
sensible thresholds for binary classifications based on continuous metrics
(rainfall, maximum sustained winds, and distance from the storm's track).
Results and conclusions would differ somewhat with other threshold choices. We
have published code for this analysis online, allowing
other researchers can explore other threshold choices for these
analyses.

@@

To conduct tropical cyclone epidemiological studies that span multiple
communities and storms, it is critical to have consistent and comprehensive
measurements of exposure to storm hazards. Here we have created and shared a
dataset that provides these data for counties in the United States over
multiple years. Despite some limitations in these data, they provide a powerful
tool for expanding tropical cyclone epidemiology studies to more extensively
leverage existing administrative health data, allowing researchers to
investigate how these storms affect county-wide health risk.  Further, this
dataset provides hazard measurements that are comparable across communities and
storms, which allows epidemiological researchers to design studies to explore
how health risks are modified by characteristics of both the storms and the
communities that are hit. The data are given in an open-source format, along
with associated software tools, which allows them to be freely used and for
others to explore all associated code and to contribute additions through
platforms like GitHub.

Based on our analysis in this paper, these data are typically in agreement with
measurements from other sources of data available to characterize
storm-associated hazards (e.g., ground-based monitors, streamgages, post-storm
wind radii estimates).  However, researchers who are planning to use the data
should explore the analyses presented in this paper to understand the strengths
and weaknesses of the data.  Our results suggest that county-level storm
exposure is not well-characterized by the closest distance that a storm's
central track came to a county, and that exposure to one storm hazard within a
county (e.g., severe winds) does not imply exposure to other hazards (e.g.,
excessive rainfall, flooding). As a result, it is critical that researchers
consider which storm hazards are likely on the causal pathway for the outcomes
they are studying, and to characterize storm exposure in a way that captures
those specific hazards, to avoid exposure misclassification.

>>>>

Epidemiological studies can help characterize which health risks are elevated
during disasters, to what degree, and to
whom~\parencite{ibrahim2005unfortunate, noji2005disasters}.  As a result, these
studies help improve disaster preparedness and
response~\parencite{noji2005disasters}.  However, tropical cyclones are
multi-hazard events, making it complicated to measure exposure and so to
conduct multi-community, multi-year studies leveraging large administrative
datasets.  Here, we provide open-source county-level data for several tropical
cyclone exposures and explore limitations in that data.  
Further, we explore patterns in storm exposure classifications based on
different metrics, and we find that county-level tropical cyclone exposure
assessments vary substantially when using different metrics, based either on
storm hazards or on how close the storm's central track came to the community.
Our results can inform exposure assessment for future county-level studies of
the health risks and impacts associated with tropical cyclones exposure as
well as insights to inform epidemiological study design.

@@

The open-source data generally correspond well with data from other potential
sources (Figures~\ref{fig:raincomparison}\,--\,\ref{fig:floodcomparison}
and~S2), but there are some caveats. The rainfall data are generally
well-correlated with ground-based observations, but may sometimes oversmooth
very high rainfall values (Figure~\ref{fig:raincomparison}). When rainfall data
is used to create binary exposure classifications, this disagreement is
unlikely to influence results, as both data sources agree in identifying these
as storms with high rainfall, but would be important to consider for cases that
include rainfall as a continuous measurement.

The peak sustained wind estimates are based on modeled, rather than observed,
values, and while the modeled wind data generally agree well with post-analysis
maximum wind radii from \ac{HURDAT2}~\parencite{landsea2013}
(Figure~\ref{fig:windcomparison}), there were a few storms with some
discrepancies. These storms---for example, Hurricane Sandy in 2012 and
Hurricane Ike in 2008---were unusually large systems for which high winds
persisted well inland from landfall (Figures~\ref{fig:windcomparison} and~S2).

For the flooding data, we found that flood event status, as determined based on
the NOAA Storm Events listings, typically agreed with measurements from
\ac{USGS} streamgages, with a flood event more likely to be listed if a storm
elevated streamflow at streamgages across the county
(Figure~\ref{fig:floodcomparison}).  However, there are differences between the
two flooding datasets, and these highlight both the difficulty of measuring
flood exposure at the county level and inherent challenges in using data from a
storm event database for epidemiologic exposure assessment.  For example, there
was one storm in Fulton County, GA, for which there was high streamflow but not
an associated flood event listing (Hurricane Ida, 2009).  This storm occurred
in November 2009, following a month with historic rainfall and flooding in
Georgia~\parencite{shepherd2011overview}.  In this case, the flooding
associated with Ida was incorporated into an ongoing flood event listing, with
a start date well before the five-day window we used to temporally match storm
event listings with tropical cyclone tracks.

@@

While these are important caveats for the data, we selected these data sources
as among the best currently available for measuring each of these hazards
consistently and comprehensively at a multi-county, multi-year scale. In
addition to providing tropical cyclone exposure metrics for individual hazards,
this dataset and its associated software allow users not only to access
measurements for single hazards, but also to create tropical cyclone exposure
profiles based on multiple hazards or to craft exposure indices that combine
hazard metrics~\parencite{chakraborty2005population, peduzzi2009assessing}.
This functionality can be critical, as different hazards of tropical cyclones can
act synergistically in causing impacts~\parencite{smith2009}.

@@

We found that tropical cyclones tended to bring different hazards to different
counties, and so agreement was typically low between distance-based
tropical cyclone exposure assessment and each of the single hazard-specific
exposures, as well as between pairs of hazard-specific metrics
(Figures~\ref{fig:ivanexposure}\,--\,\ref{fig:jaccard} and S4).  These
findings align with previous results from atmospheric science and related
fields on the characteristics of tropical cyclones. While tropical cyclone
rainfall and windspeed can be well-correlated when the tropical cyclone is over
water~\parencite{cerveny2000}, this relationship often weakens once the
hurricane has made landfall~\parencite{jiang2008}.  Fast-moving tropical
cyclones bring higher risks of dangerous winds inland~\parencite{kruk2010},
while slow-moving tropical cyclones are likely to bring more
rain~\parencite{rappaport2000} and cause more damage because of sustained
hazardous conditions~\parencite{rezapour2014}. Further, while the likelihood
and extent of flooding during a tropical cyclone is related to the tropical
cyclone's rainfall, it is also driven by factors like top soil saturation and
the structure of the water basin's drainage network~\parencite{chen2015,
rees2001}.

Based on our results, the use of a distance-based metric to assess exposure to any of
these hazards, or the use of measurements from one hazard as a proxy for
exposure to any of the other hazards considered, would often introduce exposure
misclassification. This conclusion reinforces similar findings from a study of
Florida's 2004 storm season~\parencite{grabich2015measuring}.  For some
studies, such exposure misclassification might plausibly be differential.  For
example, tropical cyclone wind exposures tend to be concentrated in counties
near the coast, while tropical cyclone wind exposures sometimes extend well
inland.  If the etiologically-relevant exposure for a health outcome is extreme
rainfall but exposure is classified based on measurements of wind, the
probability of being misclassified as unexposed would be higher in inland
counties, while the probability of being misclassified as exposed would be
higher in coastal counties. If coastal counties differ from inland counties in
either the outcome of interest or in factors associated with risk of that
outcome, exposure misclassification would be
differential~\parencite{savitz2016interpreting}.  Such differential exposure
misclassification could bias estimates of tropical cyclone effects either
towards the null (estimating a lower or null association compared to the true
association that exists) or away from the null (estimating a larger association
than actually exists)~\parencite{savitz2016interpreting, armstrong1998effect}.

We did, however, find a small set of tropical cyclones for which for which
agreement was high across several single-hazard exposure assessments (e.g.,
Floyd in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996; Ernesto in~2006
(Figure~\ref{fig:jaccard})).  Hurricanes Floyd in~1999 and Irene in~2011 both
made their first \ac{US} landfall in North Carolina at minor hurricane
windspeeds (Category~2 and~1, respectively) and then skimmed the eastern coast
of the \ac{US} north through New England, bringing substantial rainfall to much
of the eastern coast from North Carolina north and causing extensive inland
flooding in North Carolina (Floyd) and New England
(Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.  Hurricanes Hannah
in~2008, Bertha in~1996, and Ernesto in~2006 also followed the eastern
coastline. For these storms, the tropical cyclones' persistent proximity to
water may have helped maintain wind speeds in similar patterns to rain and
distance-based exposure assessments, resulting in more similarities across
exposure assessments compared to other tropical cyclones.  For these storms, it
may be possible to assess exposure to multiple hazards of the storm using a
single metric, in some cases even a proxy like the distance between the county
and the storm's track.  However, for these storms it may be difficult to
untangle the contribution of each hazard to the overall effect of the storm,
given that several hazards have similar geographical patterns.

@@

The dataset presented here does have several limitations, in addition to the
caveats previously discussed. First, the dataset is not comprehensive of all
tropical cyclone hazards. For example, coastal counties can experience
dangerous storm surge, which is not specifically covered in this dataset
(although some resulting coastal flooding is captured). We are exploring ways
to include this in future versions of the dataset, but to date we have
focused on exposures that could affect any county, whether inland or coastal.
Second, this data is aggregated to the county level. Many health outcome
datasets are aggregated at this level, but some may be aggregated at a finer
spatial resolution (e.g., Census tract or ZIP code) or unaggregated (i.e.,
point locations for each outcome). We have published the wind model used to
create this dataset as its own open-source R package
\parencite{stormwindmodel}, and it can be used to model storm-associated winds
at a finer spatial resolution; however, measurements of other hazards cannot
similarly be re-scaled through tools we provide. Next, we provide these data
and associated software tools through R packages, and so some experience in the
R programming language is required to make full use of them. However, R is
currently a popular programming language for environmental epidemiology,
allowing the data to reach a large audience, and we are exploring options to
create a web application using the Shiny platform to allow broader web-based
access of the data \parencite{shiny2019}.  Finally, to assess patterns and
agreement for binary exposure classifications, we have chosen one set of
sensible thresholds for binary classifications based on continuous metrics
(rainfall, maximum sustained surface wind, and distance from the storm's track).
Results and conclusions would differ somewhat with other threshold choices. We
have published code for this analysis online, allowing
other researchers to explore other threshold choices for these
analyses.

@@

To conduct tropical cyclone epidemiological studies that span multiple
communities and storms, it is critical to have consistent and comprehensive
measurements of exposure to storm hazards. Here we have created and shared a
dataset that provides these data for counties in the United States over
multiple years. Despite some limitations in these data, they provide a powerful
tool for expanding tropical cyclone epidemiology studies to more extensively
leverage existing administrative health data, allowing researchers to
investigate how these storms affect county-wide health risk.  Further, this
dataset provides hazard measurements that are comparable across communities and
storms, allowing epidemiological researchers to design studies to explore
how health risks are modified by characteristics of both the storms and the
communities that are hit. The data are given in an open-source format, along
with associated software tools, which allows them to be freely used and for
others to explore all associated code and to contribute additions through
platforms like GitHub.

Based on our analysis in this paper, these data are typically in agreement with
measurements from other sources of data available to characterize
storm-associated hazards (e.g., ground-based monitors, streamgages, post-storm
wind radii estimates).  However, researchers who are planning to use the data
should explore the analyses presented in this paper to understand the strengths
and weaknesses of the data.  Further, our results indicate that county-level storm
exposure is not well-characterized by the closest distance that a storm's
central track came to a county, and that exposure to one storm hazard within a
county (e.g., severe winds) does not imply exposure to other hazards (e.g.,
excessive rainfall, flooding). As a result, it is critical that researchers
consider which storm hazards are likely on the causal pathway for the outcomes
they are studying, and to characterize storm exposure in a way that captures
those specific hazards, to avoid exposure misclassification.

### 2019-11-28 writing/ehp_figures.tex
Work some more on abstract and small wording
Continue work on text

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.8\linewidth]{windcomparison}
	\caption{Comparison of two sources of wind exposure estimates: 
	(1) modeled maximum sustained surface wind speeds 
	and (2) estimates from \ac{HURDAT2}'s wind speed radii.
	Each point represents a storm, with the x axis giving the percent of
	counties classified in the same category of maximum sustained surface
	wind speeds ($<$34~\si{\knot}; 34\,--\,49.9~\si{\knot};
	50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}) by both sources of 
	data. The color of each point gives the number of study counties that were exposed to
	maximum sustained surface wind speeds of at least 34~\si{\knot} (based
	on modeled wind speeds). Estimates are shown for study storms since
	2004, the earliest year for which post-storm reanalysis wind speed
	radii estimates are routinely available in \ac{HURDAT2}, and for which at
	least one study county had a sustained wind speed of $\ge$34~\si{\knot}
	based on the \ac{HURDAT2} wind radii.
	}
\label{fig:windcomparison}
\end{figure}

@@

\begin{figure}%[tbhp] 
	\centering 
	\includegraphics[width = 0.8\linewidth]{jaccard_heatmap.pdf}
	\caption{Agreement between exposure
	classifications based on different exposure metrics for all storms
	between~1996 and~2011 for which at least~100 counties were exposed
	based on at least one metric.  Each row shows one storm, and the color
	of each cell shows the measured Jaccard index for each pair of exposure
	metrics (proportion of counties classified as exposed by both metrics
	out of counties classified as exposed by either metric). For 
	Grace in 2003 and Ida in 2009, there were no county exposures for
	either the tornado-based metric or the wind-based metric (indicated 
	by gray squares). Colors to the right of the
	heatmap show the number of exposed counties
	based on any of the metrics. Storms are displayed
	within clusters that have similar patterns based on hierarchical 
	clustering (complete link method).}
\label{fig:jaccard} 
\end{figure}

>>>>

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.8\linewidth]{windcomparison}
	\caption{Comparison of two sources of wind exposure estimates: 
	(1) modeled peak sustained surface wind 
	and (2) estimates based on \ac{HURDAT2}'s wind speed radii.
	Each point represents a storm, with the x axis giving the percent of
	counties classified in the same category of peak sustained surface
	wind ($<$34~kt; 34\,--\,49.9~kt;
	50\,--\,63.9~kt; $\ge$64~kt) by both sources of 
	data. The color of each point gives the number of study counties that were exposed to
	peak sustained surface wind of at least 34~kt (based
	on modeled wind). Estimates are shown for study storms since
	2004, the earliest year for which post-storm reanalysis wind speed
	radii estimates are routinely available in \ac{HURDAT2}, and for which at
	least one study county had a peak sustained wind of $\ge$34~\si{\knot}
	based on the \ac{HURDAT2} wind radii.
	}
\label{fig:windcomparison}
\end{figure}

@@

\begin{figure}%[tbhp] 
	\centering 
	\includegraphics[width = 0.8\linewidth]{jaccard_heatmap.pdf}
	\caption{Agreement between exposure
	classifications based on different exposure metrics for all storms
	between~1996 and~2011 for which at least~100 counties were exposed
	based on at least one metric.  Each row shows one storm, and the color
	of each cell shows the measured Jaccard index for each pair of exposure
	metrics (proportion of counties classified as exposed by both metrics
	out of counties classified as exposed by either metric). For 
	Grace in 2003 and Ida in 2009, there were no county exposures for
	either the tornado-based metric or the wind-based metric (indicated 
	by gray squares). Colors to the right of the
	heatmap show the number of exposed counties
	based on any of the metrics. Storms are displayed
	within clusters that have similar patterns based on hierarchical 
	clustering.}
\label{fig:jaccard} 
\end{figure}
